Avoiding accuracy-limiting pitfalls in the study of protein-ligand interactions with isothermal titration calorimetry

### 2015-08-02 manuscript/itc-worksheet.tex

\color{red}
{\bf Error propagation.}
The general rule for random error propagation for a quantity $f(x,y,z,\ldots)$ dependent on \emph{independent} measurements $x,y,z,\ldots$ gives squared standard error $s_f$ in $f$ as~\cite{taylor-error-propagation},
\begin{equation}
s_f^2 = \left(\frac{\partial f}{\partial x}\right)^2 s_x^2 + \left(\frac{\partial f}{\partial y}\right)^2 s_y^2 + \left(\frac{\partial f}{\partial z}\right)^2 s_z^2 + \cdots \label{equation:propagation-of-error}
\end{equation}
where $s_x$, $s_y$, and $s_z$ are the standard errors of the corresponding measurements.
This form, based on a Taylor expansion of the function $f$, can be extended to any number of contributing terms (e.g., multiple solution preparation steps)~\cite{taylor-error-propagation}.  
If $f$ can be written $f(x,y,z,\ldots) = x^i y^j z^k \cdots$---where $i$, $j$, and $k$ are powers to which $x$, $y$, and $z$ are raised---then Eq.~\ref{equation:propagation-of-error} assumes the simple form for the relative error $(s_f/f)$,
\begin{equation}
\left(\frac{s_f}{f}\right)^2 = i^2 \left(\frac{s_x}{x}\right)^2 + j^2 \left(\frac{s_y}{y}\right)^2 + k^2 \left(\frac{s_y}{z}\right)^2 + \cdots \label{equation:relative-error} .
\end{equation}
Often, a single term in Eq.~\ref{equation:relative-error} will dominate, and the relative error is essentially identical to this contribution.
For example, in the calculation of our titrant concentration from mass $m$ and volume $v$, $c = m/v$ with $(s_m/m)$ = 1\% and $(s_v/v)$ = 0.2\%, then the RSE in concentration is $(s_c/c) \approx 1\%$.
We will utilize this scheme to propagate error throughout our experiment, as well as to incorporate these errors alongside the least squares fit error in thermodynamic parameters produced by standard calorimetry analysis software.
To simplify this process for typical applications, the provided spreadsheet performs much of this error propagation automatically.
\color{black}

@@

{\bf Syringe backlash and the first-injection anomaly.}
Our GE/MicroCal VP-ITC instrument has a syringe assembly that utilizes a worm gear which, after the recommended purge-refill process, will cause a titrant shortfall in the first injection unless a ``down syringe'' command is issued prior to loading the syringe into the sample cell~\cite{tellinghuisen:anal-biochem:2004:first-injection-anomaly}; we therefore executed a 10~$\mu$L ``down syringe'' command immediately after the purge-refill cycle.
Because the instrument can take a substantial (but variable) period of time to stabilize at the desired experimental temperature after loading the syringe, significant ($>$ 0.1 $\mu$L) diffusive loss can also contribute to a first injection shortfall.
We therefore programmed an initial 1 $\mu$L ``throwaway injection'' to avoid the need to correct for diffusive titrant loss during the first 10 $\mu$L injection.
The contribution from this initial 1 $\mu$L ``throwaway'' injection was excluded from the fitting procedure during analysis.
\color{red}
Note that even though we exclude this heat from the analysis, we still need the syringe ``down'' command to ensure that the correct amount of titrant enters the cell.
\color{black}

@@

The protein concentration was determined spectrophotometrically via absorbance at 280 nm on a NanoDrop ND-1000.
\color{red}
The NanoDrop (and similar instruments) utilize small sample volumes (3 $\mu$L was used here) and dynamic selection of among path lengths (between 0.2 mm and 1 mm for the ND-1000) to facilitate direct determination of typical protein concentrations without dilution.
Here the measured absorbance of 1.18$\pm$0.02 at 1 mm path [henceforth written 1.18(2)] length yielded a protein concentration of 235$\pm$4 $\mu$M using the known molar absorptivity $\epsilon_\mathrm{280\:nm} = 50070(25)$ M$^{-1}$ cm$^{-1}$~\cite{myszka:2003:j-biomol-tech:abrf-mirg02}.
The sample was then diluted to [M]$_0$ $\approx$ 10 $\mu$M using the purity-corrected post-dialysis concentration. 
Note that high precision is not generally required for protein concentration determination unless the binding stoichiometry is unknown, as the site parameter $n$ absorbs errors in [M]$_0$ and $V_0$ in standard least-squares data analysis~\cite{tellinghuisen:2012:anal-biochem:designing-itc-experiments}.
\color{black}

@@

To load the VP-ITC syringe, we require $\sim$2.1 mL of our titrant\footnote{A smaller 700 $\mu$L filling tube is also available.}.  
For our chosen [X]$_s$ $\sim$ 720 $\mu$M, this requires only 0.3 mg of CBS, but given the precision of the analytical balance used for this step (Mettler-Toledo AB204, readability $\pm$0.1 mg), this would yield 33\% uncertainty in [X]$_s$, and hence the final relative errors in $K_a$ and $\Delta H^\circ$ would be \emph{at least} this large.  
To reduce the mass uncertainty to 1\%, we must weigh out at least 10 mg.  
Since the solubility of CBS in water is only 453 mg/L at room temperature (which corresponds to a 2~250 $\mu$M solution), we need a volume of at least 22 mL to dissolve 10 mg.
Using a 25 mL Class A volumetric flask or pipette (rated $\pm$0.05 mL) would allow us to attain the desired 1\% precision.  
On the other hand, graduated cylinders and serological pipettes with 25 mL capacity often possess a precision of only $\pm$0.5 ml, which would raise the uncertainty in [X]$_s$ to 2\%.  
Here, we found it convenient to employ multiple liquid transfers with a Gilson P5000 5 mL pipette, which has a stated reliability of $\pm$0.03 mL at full capacity\footnote{\color{red}For pipettes, the stated systematic error $\delta$ is generally larger than the imprecision.  If the same pipette is used to deliver \emph{multiple} aliquots, the uncertainty for the total volume transferred should be estimated from summing the systematic error $\delta$ for each transfer.  Since the systematic error is generally unknown and we utilize each pipette once, we use the stated systematic error to estimate the uncertainty in transferred volume, assuming it behaves like random error.  For example, if 25 mL is transferred in five transfers of 5 mL using a P5000 ($\delta = 30$ $\mu$L), the uncertainty is $(5)(0.030) = 0.15$ mL.}.

We chose to prepare a 1~500 $\mu$M CBS stock solution as a compromise between ensuring complete solubility of CBS (solubility 2 250 $\mu$M in water) and minimizing buffer use (preparing a solution of $\sim$720 $\mu$M directly with 10 $\mu$g CBS would have doubled the quantity of buffer required).
\color{red}
To do this, we added 10.0(1) mg CBS to 32.1(2) mL PBS dialysate and vortexed to ensure the compound was completely dissolved, yielding 32.2(2) mL of a 1.50(2) mM CBS stock solution.  
\color{black}

To ensure sufficient $\sim$718.29 $\mu$M titrant to allow for a ligand-into-buffer blank titration and additional experimental replications if needed, we planned to prepare 9 mL of titrant solution.
This is more than necessary, as minimum of 700 $\mu$L/experiment is required for the VP-ITC if the low-volume syringe loading tube is utilized.
\color{red}
Using the Gilson P5000, we then added 4.309(12) mL CBS stock to 4.691(12) mL PBS to obtain a 717(9) $\mu$M CBS titrant (1.2\% RSE).  
\color{black}
Error propagation was performed automatically by the spreadsheet (Figure \ref{figure:spreadsheet}).

@@

{\bf Data analysis.}
\color{red}
The titration dataset (Figure~\ref{figure:enthalpogram}) was analyzed using Origin 7.0 (OriginLab Corp.) after subtracting heats obtained from a separate ligand-into-buffer “blank” titration utilizing the same protocol (Supplementary Figure~\ref{figure:blank-heats}).
Here, the blank heats were small and uniform, of the same magnitude as water-into-water injections.  
The least-squares (LS) fit of the thermodynamic parameters to the integrated injection heats are shown in the caption of Figure~\ref{figure:enthalpogram}. 
Note that, since the stoichiometry is known to be 1:1, the site parameter $n$ absorbs errors in [M]$_0$ and the cell volume V$_0$; if the actual concentration of active macromolecule is of interest, these quantities will require more precision~\cite{tellinghuisen:anal-biochem:2004:volume-errors-in-itc}.
While $\Delta H^\circ$ is rather insensitive to errors in the stated cell volume V$_0$ as a result, those errors can have a substantial effect on $K_a$, so careful calibration of V$_0$ using standard reactions~(e.g.~\cite{tellinghuisen:biophys-chem:2004:barium-crown-ether-itc,tellinghuisen:anal-biochem:2007:itc-calibration-nacl}) is advised if highly accurate $K_a$ is sought~\cite{tellinghuisen:anal-biochem:2004:volume-errors-in-itc}.
Raw and processed datasets are provided with this work as Supplementary Material.
\color{black}

@@

\color{red}
Since $\Delta G^\circ$ logarithmically depends on $K_a$ through the relation $\Delta G^\circ = -RT \ln [K_a C_\circ]$, the uncertainty in $\Delta G^\circ$ computed using Eq.~\ref{equation:relative-error-in-thermodynamic-parameters}---where $s_{\Delta G^\circ} = RT (s_{K_a} / K_a)$ = 0.02 kcal/mol---is much smaller than that in $\Delta H^\circ$ (0.15 kcal/mol).
If the entropic contribution to binding, $-T \Delta S^\circ = \Delta G^\circ - \Delta H^\circ$ is of interest, its uncertainty can similarly be obtained from Eq.~\ref{equation:propagation-of-error}, and found to be of the same magnitude as that in $\Delta H^\circ$ (0.15 kcal/mol)\footnote{Because $\Delta H^\circ$ and $K_a$ (hence $\Delta G^\circ$ are obtained from the same fit---and hence are correlated---cross-terms of the form $2 (\partial f/\partial x)(\partial f/\partial y) s_{xy}$ with $x \equiv \Delta G^\circ$ and $y \equiv - \Delta H^\circ$ must be added to Eq.~\ref{equation:propagation-of-error}, but because the uncertainty in $\Delta H^\circ$ is an order of magnitude larger than that in $\Delta G^\circ$, it still dominates the overall uncertainty even if these correlation terms are included.}.  
\color{black}

@@

{\bf Discussion.}
\color{red}
Note that our excess uncertainty comes directly from the uncertainty in the prepared titrant concentration $[X]_s$.
Had we chosen to use much less than 10 mg of compound, or utilized low-precision volume transfer devices (such as serological pipettes), we could have easily raised this contribution to 10\% or more, which would then dominate our apparent LS uncertainties.
\color{black}
Although the absolute error in $\Delta G^\circ$ would remain small ($\sim$0.04 kcal/mol), the absolute error in $\Delta H^\circ$ would be large ($\sim$1.1 kcal/mol), making the error in $-T \Delta S^\circ$ comparable in magnitude.  
This can have important consequences in trying to ascribe significance to differences in entropy-enthalpy compensation within a congeneric series, especially when differences in $\Delta G^\circ$ are small~\cite{freire:chem-biol-drug-des:2007:compensation,freire:drug-discovery-today:2008:entropy-enthalpy,freire:nat-rev-drug-discovery:2010:itc-lead-optimization,chodera-mobley:annu-rev-biophys:2013:entropy-enthalpy-compensation}.

\color{red}
We recall that the reported errors in $\Delta H^\circ$ (and hence $T \Delta S^\circ$) for the ABRF-MIRG'02 study were as much as two orders of magnitude smaller than the actual error deduced from variation among independent measurements.
If indeed concentration errors were at fault, simply repeating the experiment with the same solutions would not have revealed any problem~\cite{tellinghuisen-chodera:2011:anal-biochem:systematic-itc-errors,chodera-mobley:annu-rev-biophys:2013:entropy-enthalpy-compensation}.
\color{black}

@@

\end{centering}
\caption{{\bf Spreadsheet for this experiment showing automated propagation of error.}  
This spreadsheet and blank templates is available for download in multiple formats at \url{http://github.org/choderalab/itc-worksheet}.
\color{red}
Note that some quantities are recorded to greater precision than experimental uncertainty in the spreadsheet by virtue of having been recorded directly from the instrument.
These quantities are always written in the text with appropriate attention to significant figures---that is, only the largest significant figure of the uncertainty is recorded, and the value it is attached to is truncated to that decimal place.
\color{black}
\label{figure:spreadsheet}}
\end{figure*}
%
%\subsection{Propagation of pipetting error}
%
%{\color{red}[JDC: This section should have more information on propagation of error for pipetting steps and a summary of common pipette errors.  I'm still trying to figure out if there is a way to have this done automatically in the spreadsheet, such as having the experimenter list the number of volume transfers of a particular pipette class.]}

>>>>

{\bf Error propagation.}
The general rule for random error propagation for a quantity $f(x,y,z,\ldots)$ dependent on \emph{independent} measurements $x,y,z,\ldots$ gives squared standard error $s_f$ in $f$ as~\cite{taylor-error-propagation},
\begin{equation}
s_f^2 = \left(\frac{\partial f}{\partial x}\right)^2 s_x^2 + \left(\frac{\partial f}{\partial y}\right)^2 s_y^2 + \left(\frac{\partial f}{\partial z}\right)^2 s_z^2 + \cdots \label{equation:propagation-of-error}
\end{equation}
where $s_x$, $s_y$, and $s_z$ are the standard errors of the corresponding measurements.
This form, based on a Taylor expansion of the function $f$, can be extended to any number of contributing terms (e.g., multiple solution preparation steps)~\cite{taylor-error-propagation}.  
If $f$ can be written $f(x,y,z,\ldots) = x^i y^j z^k \cdots$---where $i$, $j$, and $k$ are powers to which $x$, $y$, and $z$ are raised---then Eq.~\ref{equation:propagation-of-error} assumes the simple form for the relative error $(s_f/f)$,
\begin{equation}
\left(\frac{s_f}{f}\right)^2 = i^2 \left(\frac{s_x}{x}\right)^2 + j^2 \left(\frac{s_y}{y}\right)^2 + k^2 \left(\frac{s_y}{z}\right)^2 + \cdots \label{equation:relative-error} .
\end{equation}
Often, a single term in Eq.~\ref{equation:relative-error} will dominate, and the relative error is essentially identical to this contribution.
For example, in the calculation of our titrant concentration from mass $m$ and volume $v$, $c = m/v$ with $(s_m/m)$ = 1\% and $(s_v/v)$ = 0.2\%, then the RSE in concentration is $(s_c/c) \approx 1\%$.
We will utilize this scheme to propagate error throughout our experiment, as well as to incorporate these errors alongside the least squares fit error in thermodynamic parameters produced by standard calorimetry analysis software.
To simplify this process for typical applications, the provided spreadsheet performs much of this error propagation automatically.

@@

{\bf Syringe backlash and the first-injection anomaly.}
Our GE/MicroCal VP-ITC instrument has a syringe assembly that utilizes a worm gear which, after the recommended purge-refill process, will cause a titrant shortfall in the first injection unless a ``down syringe'' command is issued prior to loading the syringe into the sample cell~\cite{tellinghuisen:anal-biochem:2004:first-injection-anomaly}; we therefore executed a 10~$\mu$L ``down syringe'' command immediately after the purge-refill cycle.
Because the instrument can take a substantial (but variable) period of time to stabilize at the desired experimental temperature after loading the syringe, significant ($>$ 0.1 $\mu$L) diffusive loss can also contribute to a first injection shortfall.
We therefore programmed an initial 1 $\mu$L ``throwaway injection'' to avoid the need to correct for diffusive titrant loss during the first 10 $\mu$L injection.
The contribution from this initial 1 $\mu$L ``throwaway'' injection was excluded from the fitting procedure during analysis.
Note that even though we exclude this heat from the analysis, we still need the syringe ``down'' command to ensure that the correct amount of titrant enters the cell.

@@

The protein concentration was determined spectrophotometrically via absorbance at 280 nm on a NanoDrop ND-1000.
The NanoDrop (and similar instruments) utilize small sample volumes (3 $\mu$L was used here) and dynamic selection of among path lengths (between 0.2 mm and 1 mm for the ND-1000) to facilitate direct determination of typical protein concentrations without dilution.
Here the measured absorbance of 1.18$\pm$0.02 at 1 mm path [henceforth written 1.18(2)] length yielded a protein concentration of 235$\pm$4 $\mu$M using the known molar absorptivity $\epsilon_\mathrm{280\:nm} = 50070(25)$ M$^{-1}$ cm$^{-1}$~\cite{myszka:2003:j-biomol-tech:abrf-mirg02}.
The sample was then diluted to [M]$_0$ $\approx$ 10 $\mu$M using the purity-corrected post-dialysis concentration. 
Note that high precision is not generally required for protein concentration determination unless the binding stoichiometry is unknown, as the site parameter $n$ absorbs errors in [M]$_0$ and $V_0$ in standard least-squares data analysis~\cite{tellinghuisen:2012:anal-biochem:designing-itc-experiments}.

@@

To load the VP-ITC syringe, we require $\sim$2.1 mL of our titrant\footnote{A smaller 700 $\mu$L filling tube is also available.}.  
For our chosen [X]$_s$ $\sim$ 720 $\mu$M, this requires only 0.3 mg of CBS, but given the precision of the analytical balance used for this step (Mettler-Toledo AB204, readability $\pm$0.1 mg), this would yield 33\% uncertainty in [X]$_s$, and hence the final relative errors in $K_a$ and $\Delta H^\circ$ would be \emph{at least} this large.  
To reduce the mass uncertainty to 1\%, we must weigh out at least 10 mg.  
Since the solubility of CBS in water is only 453 mg/L at room temperature (which corresponds to a 2~250 $\mu$M solution), we need a volume of at least 22 mL to dissolve 10 mg.
Using a 25 mL Class A volumetric flask or pipette (rated $\pm$0.05 mL) would allow us to attain the desired 1\% precision.  
On the other hand, graduated cylinders and serological pipettes with 25 mL capacity often possess a precision of only $\pm$0.5 ml, which would raise the uncertainty in [X]$_s$ to 2\%.  
Here, we found it convenient to employ multiple liquid transfers with a Gilson P5000 5 mL pipette, which has a stated reliability of $\pm$0.03 mL at full capacity\footnote{For pipettes, the stated systematic error $\delta$ is generally larger than the imprecision.  If the same pipette is used to deliver \emph{multiple} aliquots, the uncertainty for the total volume transferred should be estimated from summing the systematic error $\delta$ for each transfer.  Since the systematic error is generally unknown and we utilize each pipette once, we use the stated systematic error to estimate the uncertainty in transferred volume, assuming it behaves like random error.  For example, if 25 mL is transferred in five transfers of 5 mL using a P5000 ($\delta = 30$ $\mu$L), the uncertainty is $(5)(0.030) = 0.15$ mL.}.

We chose to prepare a 1~500 $\mu$M CBS stock solution as a compromise between ensuring complete solubility of CBS (solubility 2 250 $\mu$M in water) and minimizing buffer use (preparing a solution of $\sim$720 $\mu$M directly with 10 $\mu$g CBS would have doubled the quantity of buffer required).
To do this, we added 10.0(1) mg CBS to 32.1(2) mL PBS dialysate and vortexed to ensure the compound was completely dissolved, yielding 32.2(2) mL of a 1.50(2) mM CBS stock solution.

To ensure sufficient $\sim$718.29 $\mu$M titrant to allow for a ligand-into-buffer blank titration and additional experimental replications if needed, we planned to prepare 9 mL of titrant solution.
This is more than necessary, as minimum of 700 $\mu$L/experiment is required for the VP-ITC if the low-volume syringe loading tube is utilized.
Using the Gilson P5000, we then added 4.309(12) mL CBS stock to 4.691(12) mL PBS to obtain a 717(9) $\mu$M CBS titrant (1.2\% RSE).  
Error propagation was performed automatically by the spreadsheet (Figure \ref{figure:spreadsheet}).

@@

{\bf Data analysis.}
The titration dataset (Figure~\ref{figure:enthalpogram}) was analyzed using Origin 7.0 (OriginLab Corp.) after subtracting heats obtained from a separate ligand-into-buffer “blank” titration utilizing the same protocol (Supplementary Figure~\ref{figure:blank-heats}).
Here, the blank heats were small and uniform, of the same magnitude as water-into-water injections.  
The least-squares (LS) fit of the thermodynamic parameters to the integrated injection heats are shown in the caption of Figure~\ref{figure:enthalpogram}. 
Note that, since the stoichiometry is known to be 1:1, the site parameter $n$ absorbs errors in [M]$_0$ and the cell volume V$_0$; if the actual concentration of active macromolecule is of interest, these quantities will require more precision~\cite{tellinghuisen:anal-biochem:2004:volume-errors-in-itc}.
While $\Delta H^\circ$ is rather insensitive to errors in the stated cell volume V$_0$ as a result, those errors can have a substantial effect on $K_a$, so careful calibration of V$_0$ using standard reactions~(e.g.~\cite{tellinghuisen:biophys-chem:2004:barium-crown-ether-itc,tellinghuisen:anal-biochem:2007:itc-calibration-nacl}) is advised if highly accurate $K_a$ is sought~\cite{tellinghuisen:anal-biochem:2004:volume-errors-in-itc}.
Raw and processed datasets are provided with this work as Supplementary Material.

@@

Since $\Delta G^\circ$ logarithmically depends on $K_a$ through the relation $\Delta G^\circ = -RT \ln [K_a C_\circ]$, the uncertainty in $\Delta G^\circ$ computed using Eq.~\ref{equation:relative-error-in-thermodynamic-parameters}---where $s_{\Delta G^\circ} = RT (s_{K_a} / K_a)$ = 0.02 kcal/mol---is much smaller than that in $\Delta H^\circ$ (0.15 kcal/mol).
If the entropic contribution to binding, $-T \Delta S^\circ = \Delta G^\circ - \Delta H^\circ$ is of interest, its uncertainty can similarly be obtained from Eq.~\ref{equation:propagation-of-error}, and found to be of the same magnitude as that in $\Delta H^\circ$ (0.15 kcal/mol)\footnote{Because $\Delta H^\circ$ and $K_a$ (hence $\Delta G^\circ$ are obtained from the same fit---and hence are correlated---cross-terms of the form $2 (\partial f/\partial x)(\partial f/\partial y) s_{xy}$ with $x \equiv \Delta G^\circ$ and $y \equiv - \Delta H^\circ$ must be added to Eq.~\ref{equation:propagation-of-error}, but because the uncertainty in $\Delta H^\circ$ is an order of magnitude larger than that in $\Delta G^\circ$, it still dominates the overall uncertainty even if these correlation terms are included.}.

@@

{\bf Discussion.}
Note that our excess uncertainty comes directly from the uncertainty in the prepared titrant concentration $[X]_s$.
Had we chosen to use much less than 10 mg of compound, or utilized low-precision volume transfer devices (such as serological pipettes), we could have easily raised this contribution to 10\% or more, which would then dominate our apparent LS uncertainties.
Although the absolute error in $\Delta G^\circ$ would remain small ($\sim$0.04 kcal/mol), the absolute error in $\Delta H^\circ$ would be large ($\sim$1.1 kcal/mol), making the error in $-T \Delta S^\circ$ comparable in magnitude.  
This can have important consequences in trying to ascribe significance to differences in entropy-enthalpy compensation within a congeneric series, especially when differences in $\Delta G^\circ$ are small~\cite{freire:chem-biol-drug-des:2007:compensation,freire:drug-discovery-today:2008:entropy-enthalpy,freire:nat-rev-drug-discovery:2010:itc-lead-optimization,chodera-mobley:annu-rev-biophys:2013:entropy-enthalpy-compensation}.

We recall that the reported errors in $\Delta H^\circ$ (and hence $T \Delta S^\circ$) for the ABRF-MIRG'02 study were as much as two orders of magnitude smaller than the actual error deduced from variation among independent measurements.
If indeed concentration errors were at fault, simply repeating the experiment with the same solutions would not have revealed any problem~\cite{tellinghuisen-chodera:2011:anal-biochem:systematic-itc-errors,chodera-mobley:annu-rev-biophys:2013:entropy-enthalpy-compensation}.

@@

\end{centering}
\caption{{\bf Spreadsheet for this experiment showing automated propagation of error.}  
This spreadsheet and blank templates is available for download in multiple formats at \url{http://github.org/choderalab/itc-worksheet}.
Note that some quantities are recorded to greater precision than experimental uncertainty in the spreadsheet by virtue of having been recorded directly from the instrument.
These quantities are always written in the text with appropriate attention to significant figures---that is, only the largest significant figure of the uncertainty is recorded, and the value it is attached to is truncated to that decimal place.
\label{figure:spreadsheet}}
\end{figure*}
%
%\subsection{Propagation of pipetting error}
%
%{\color{red}[JDC: This section should have more information on propagation of error for pipetting steps and a summary of common pipette errors.  I'm still trying to figure out if there is a way to have this done automatically in the spreadsheet, such as having the experimenter list the number of volume transfers of a particular pipette class.]}
